### 2021-05-17 project_ashish_monograph/CMS.tex
references added

\onehalfspacing Compact Muon Solenoid (CMS) is one of the major experiment at CERN's Large Hadron Collider (LHC) that have been built to search for new physics. CMS is designed to detect new particles produced in the LHC's high energy proton-proton and heavy-ion (Lead, Xenon) collisions. CMS will also measure the properties (mass, decay width, coupling) of already discovered particles like quarks and leptons with unprecedented precision and will be looking to discover completely new, beyond standard model of particle physics phenomena which will help us to get a better understanding of our Universe and may confirm or rule out multiverse hypothesis. This experiment has already confirmed the discovery of the most important particle in SM, the Higgs boson which gives mass to all bosons and fermions via electroweak symmetry breaking mechanism. It has also detected the decay of Higgs boson to a pair of photons, W bosons, bottom quarks and muons.  \\

@@

The CMS Run 2 pixel detector consists of four concentric barrel layers (L1-L4) at radii of 29, 68, 109 and 160 mm and three disks (D1-D3) on each end at distances of 291, 396, and 516 mm from the center of the detector. It is built from 1856 segmented silicon sensor modules, where 1184 modules are used in the barrel pixel detector (BPIX) and 672 modules are used for the forward disks (FPIX). \\

>>>>

\onehalfspacing Compact Muon Solenoid (CMS) is one of the major experiment at CERN's Large Hadron Collider (LHC) that have been built to search for new physics. CMS is designed to detect new particles produced in the LHC's high energy proton-proton and heavy-ion (Lead, Xenon) collisions. CMS will also measure the properties (mass, decay width, coupling) of already discovered particles like quarks and leptons with unprecedented precision and will be looking to discover completely new, beyond standard model of particle physics phenomena which will help us to get a better understanding of our Universe and may confirm or rule out multiverse hypothesis \cite{Chatrchyan:1129810}. This experiment has already confirmed the discovery of the most important particle in SM, the Higgs boson which gives mass to all bosons and fermions via electroweak symmetry breaking mechanism \cite{Chatrchyan:2012xdj}. It has also detected the decay of Higgs boson to a pair of photons, W bosons, bottom quarks and muons.  \\

@@

The CMS Run 2 pixel detector consists of four concentric barrel layers (L1-L4) at radii of 29, 68, 109 and 160 mm and three disks (D1-D3) on each end at distances of 291, 396, and 516 mm from the center of the detector. It is built from 1856 segmented silicon sensor modules, where 1184 modules are used in the barrel pixel detector (BPIX) and 672 modules are used for the forward disks (FPIX) \cite{TrackerGroupoftheCMS:2020bgg}. \\

### 2021-05-17 project_ashish_monograph/LHC.tex
progress
references added

\onehalfspacing CERN accelerator complex is a progression of machines namely Linac 2, Proton Synchrotron Booster (PSB), Proton Synchrotron (PS), Super Proton Synchrotron (SPS) and Large Hadron Collider (LHC) as shown in Fig. 1 that accelerate particles mainly protons to increasingly higher energies of 50 MeV, 1.4 GeV, 25 GeV, 450 GeV and 14 TeV respectively. Each machine is used to increase the energy of particle before the beam is injected into to the next machine. Linear accelerator (Linac2) is the starting point for protons used in various experiments at CERN. After proton energy reaches 50 MeV, protons enter PSB which further increases its energy before they are injected into PS which takes up its energy to 25 GeV. After reaching 25 GeV, protons enter SPS which further increases its energy to 450 GeV. The protons are finally transferred to two beam pipes of LHC kept at very high vacuum.  \\

The Large Hadron Collider (LHC) is the most powerful particle accelerator in the world primarily build to collide protons and heavy ions at 7, 8, 13 and 14 TeV during different LHC run period. The LHC consists of a 27 kilometre ring of superconducting magnets producing strong magnetic field which is used to guide particle beams along the LHC ring. Proton beam made up of bunches travelling in opposite directions kept in two ultrahigh vacuum tubes are accelerated close to the speed of light using time dependent electric field produced by radiofrequency cavities before they collide at four interaction points along the LHC ring where ATLAS, ALICE, CMS and LHCb detectors are placed. \\

The first proton beam was successfully maneuvered around the LHC ring in September 2008 that marked the beginning of Run 1 with proton beams colliding at 7-8 TeV center-of-mass energy and this run period ended in 2013. Run 2 started in 2015 with protons colliding at 13 TeV center-of-mass energy and continued till October 2018. Run 3 is expected to begin at the start of March 2022 and High Luminosity (HL)-LHC phase is expected to begin in mid-2027. Proton Synchroton (PS) is responsible for providing proton bunches spaced 7 metres (25 ns) apart with each bunch containing more than 100 billion protons. The maximum number of bunches reached with the beam preparation method used during Run 2 is 2556. The number of protons in the colliding bunches, time of separation between the bunches and effective area of the bunch which depend on the rms transverse bunch size along horizontal and vertical directions is one of the ways to calculate instantaneous luminosity (L_{inst}). \\

Luminosity is a key performance parameter of an accelerator as it is a measure of number of collisions. Luminosity can be categorized into two types: instantaneous and integrated luminosity. Instantaneous luminosity is defined as the number of collisions occuring per second in per centimeter square area while integrated luminosity is obtained by integrating instantaneous luminosity over a time interval. Both are very important variables as they are used to obtain the number of collision events. During Run 1, peak value of LHC instantaneous luminosity was $0.77 \times 10^{34} \: cm^{-2} s^{-1} $ with $30 \: fb^{-1}$ integrated luminosity. During Run 2, LHC instantaneous luminosity reached a peak value of $1.58 \times 10^{34}  \: cm^{-2} s^{-1}$ with $150 \: fb^{-1}$ integrated luminosity as shown in Fig. 2. At the end of Run 3, we expect to achieve an integrated luminosity  of $300 \:fb^{-1}$ while Phase 2 HL-LHC is expected to collect $3000 \:fb^{-1}$ of data as shown in Fig. 3 thereby increasing luminosity by a factor of 10 beyond the LHC's design value. \\

\begin{figure}[H]
  \centering
  \includegraphics[width=0.8\columnwidth]{./LHCcomplex.png}
  \caption{\onehalfspacing Diagram of the LHC accelerator complex located in Geneva, Switzerland. Four interaction points are shown for the ALICE, ATLAS, LHCb and CMS experiments.}
  \label{fig:LHC}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=1 \columnwidth]{./run1-2_lumi.png}
  \caption{\onehalfspacing Left: Instantaneous luminosities for Run 1 and Run 2 period. Right: Integrated luminosities for Run 1 and Run 2 period.}
  \label{fig:LHClumi}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.7 \columnwidth]{./HLLHCLumi.png}
  \caption{ \onehalfspacing Projected performance of the LHC until 2038, which shows the preliminary dates for prolonged stops (LS1, LS2, LS3, LS4, LS5) of the LHC and luminosities. Red points show instantaneous luminosity $(L_{inst})$ while the blue line shows integrated luminosity. \cite{collaborations2019report}.}
  \label{fig:LHCPlans}
\end{figure}

>>>>

\onehalfspacing CERN accelerator complex is a progression of machines namely Linac 2, Proton Synchrotron Booster (PSB), Proton Synchrotron (PS), Super Proton Synchrotron (SPS) and Large Hadron Collider (LHC) as shown in Fig. 1 that accelerate particles mainly protons to increasingly higher energies of 50 MeV, 1.4 GeV, 25 GeV, 450 GeV and 14 TeV respectively \cite {CERN-Brochure-2017-002-Eng}. Each machine is used to increase the energy of particle before the beam is injected into to the next machine. Linear accelerator (Linac2) is the starting point for protons used in various experiments at CERN. After proton energy reaches 50 MeV, protons enter PSB which further increases its energy before they are injected into PS which takes up its energy to 25 GeV. After reaching 25 GeV, protons enter SPS which further increases its energy to 450 GeV. The protons are finally transferred to two beam pipes of LHC kept at very high vacuum.  \\

The Large Hadron Collider (LHC) is the most powerful particle accelerator in the world primarily build to collide protons and heavy ions at 7, 8, 13 and 14 TeV during different LHC run period. The LHC consists of a 27 kilometre ring of superconducting magnets producing strong magnetic field which is used to guide particle beams along the LHC ring. Proton beam made up of bunches travelling in opposite directions kept in two ultrahigh vacuum tubes are accelerated close to the speed of light using time dependent electric field produced by radiofrequency cavities before they collide at four interaction points along the LHC ring where ATLAS, ALICE, CMS and LHCb detectors are placed \cite {CERN-Brochure-2017-002-Eng}. \\

The first proton beam was successfully maneuvered around the LHC ring in September 2008 that marked the beginning of Run 1 with proton beams colliding at 7-8 TeV center-of-mass energy and this run period ended in 2013 \cite{cmsrun}. Run 2 started in 2015 with protons colliding at 13 TeV center-of-mass energy and continued till October 2018 \cite{Wenninger:2668326}. Run 3 is expected to begin at the start of March 2022 and High Luminosity (HL)-LHC phase is expected to begin in mid-2027 \cite{Dainese:2019rgk}. Proton Synchroton (PS) is responsible for providing proton bunches spaced 7 metres (25 ns) apart with each bunch containing more than 100 billion protons. The maximum number of bunches reached with the beam preparation method used during Run 2 is 2556 \cite{Wenninger:2668326}. The number of protons in the colliding bunches, time of separation between the bunches and effective area of the bunch which depend on the rms transverse bunch size along horizontal and vertical directions is one of the ways to calculate instantaneous luminosity (L_{inst}). \\

Luminosity is a key performance parameter of an accelerator as it is a measure of number of collisions. Luminosity can be categorized into two types: instantaneous and integrated luminosity. Instantaneous luminosity is defined as the number of collisions occuring per second in per centimeter square area while integrated luminosity is obtained by integrating instantaneous luminosity over a time interval. Both are very important variables as they are used to obtain the number of collision events. During Run 1, peak value of LHC instantaneous luminosity was $0.77 \times 10^{34} \: cm^{-2} s^{-1} $ with $30 \: fb^{-1}$ integrated luminosity \cite{cmsrun1lumi}  . During Run 2, LHC instantaneous luminosity reached a peak value of $1.58 \times 10^{34}  \: cm^{-2} s^{-1}$ with $150 \: fb^{-1}$ integrated luminosity as shown in Fig. 2 \cite{CMS:2018elu}. At the end of Run 3, we expect to achieve an integrated luminosity  of $300 \:fb^{-1}$ while Phase 2 HL-LHC is expected to collect $3000 \:fb^{-1}$ of data as shown in Fig. 3 thereby increasing luminosity by a factor of 10 beyond the LHC's design value \cite{Dainese:2019rgk}. \\

\begin{figure}[H]
  \centering
  \includegraphics[width=0.8\columnwidth]{./LHCcomplex.png}
  \caption{\onehalfspacing Diagram of the LHC accelerator complex located in Geneva, Switzerland. Four interaction points are shown for the ALICE, ATLAS, LHCb and CMS experiments. \cite{Mobs:2684277}}
  \label{fig:LHC}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=1 \columnwidth]{./run1-2_lumi.png}
  \caption{\onehalfspacing Left: Instantaneous luminosities for Run 1 and Run 2 period. Right: Integrated luminosities for Run 1 and Run 2 period. \cite{lumi}}
  \label{fig:LHClumi}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.7 \columnwidth]{./lumi_projection.png}
  \caption{ \onehalfspacing Projected performance of the LHC until 2038, which shows the preliminary dates for prolonged stops (LS1, LS2, LS3, LS4, LS5) of the LHC and luminosities. Red points show instantaneous luminosity $(L_{inst})$ while the blue line shows integrated luminosity. \cite{collaborations2019report}.}
  \label{fig:LHCPlans}
\end{figure}

### 2021-05-17 project_ashish_monograph/PCC.tex
references added

The instantaneous luminosity per bunch crossing L is proportional to the number of proton-proton (pp) interactions per bunch crossing (pileup) and the proportionality constant is the ratio of  LHC orbit frequency f and the pp interaction cross section $\sigma_{interaction}(\sqrt{s})$ where s is the proton-proton collision center-of-mass energy. \\

@@

Specific techniques are employed for the reconstruction of clusters in the pixel detector and all reconstructed clusters are used for counting to calculate the instantaneous luminosity. Algorithm used for finding clusters starts by looking for pixels whose signal to noise ratio is more than 6 and then merge adjacent pixel. These algorithms estimates cluster position in two directions and cluster charge. A cluster is a set of adjacent pixels and only pixels above certain threshold charge are considered. Position of a cluster in X and Y direction are obtained by fitting X and Y prpjections to templates that are estimates for cluster shapes.

>>>>

The instantaneous luminosity per bunch crossing L is proportional to the number of proton-proton (pp) interactions per bunch crossing (pileup) and the proportionality constant is the ratio of  LHC orbit frequency f and the pp interaction cross section $\sigma_{interaction}(\sqrt{s})$ where s is the proton-proton collision center-of-mass energy \cite{CMS-PAS-LUM-12-001}. \\

@@

Specific techniques are employed for the reconstruction of clusters in the pixel detector and all reconstructed clusters are used for counting to calculate the instantaneous luminosity. Algorithm used for finding clusters starts by looking for pixels whose signal to noise ratio is more than 6 and then merge adjacent pixel. These algorithms estimates cluster position in two directions and cluster charge. A cluster is a set of adjacent pixels and only pixels above certain threshold charge are considered. Position of a cluster in X and Y direction are obtained by fitting X and Y projections to templates that are estimates for cluster shapes \cite{Chatrchyan:2014fea}.

### 2021-05-17 project_ashish_monograph/bkg.tex
references added

The afterglow corrections for PCC during 2017 run are in the range 2-5 $\%$ for the Type 1 corrections and 2-3 $\%$ averaged over all active bunches for the Type 2 corrections.

>>>>

The afterglow corrections for PCC during 2017 run are in the range 2-5 $\%$ for the Type 1 corrections and 2-3 $\%$ averaged over all active bunches for the Type 2 corrections \cite{Sirunyan:2759951}.

### 2021-05-17 project_ashish_monograph/upgrade.tex
references added

The High Luminosity (HL)-LHC will increase instantaneous luminosity to unprecedented value of $7.5 \times 10^{34} cm^{-2} s^{-1}$ which corresponds to 200 proton-proton collisions per bunch crossing (pileup). Run 2 pixel detector will not be able to handle the extreme radiation environment, resolve nearby particle tracks and operate properly to give a reliable estimate of the instantaneous luminosity for high pileup values. That is why it will be replaced by a new pixel detector which will be composed of three subdetectors: tracker barrel pixel detector (TBPX), tracker forward pixel detector (TFPX) and tracker endcap pixel detector (TEPX). TEPX will have better radiation tolerance, increased granularity, improved two-track separation, improved estimation of hit rate and statistical precision, extended tracking acceptance $|\eta|=4$ with Disk 4 Ring 1 operating as an independent luminometer. \\

Tracker endcap pixel detector (TEPX) consists of four double disks per side (-Z and +Z) as shown in Fig. 27 with each double disk containing five rings as shown in Fig. 28 having 20, 28, 36, 44 and 48 modules respectively. One double disk has four surfaces with +Z side containing modules with even module number in front layers (L1 $\&$ L2) and modules with odd module number in back layers (L3 $\&$ L4) from Ring 1 to Ring 4 and for Ring 5, modules with odd module number in front layers and modules with even module number in back layers as shown in Fig. 29. For -Z side, four surfaces contain modules with odd module number in front layer and modules with even module number in back layer from Ring 1 to Ring 4 and for Ring 5, modules with even module number in front layer and modules with odd module number in back layer. \\

Instantaneous luminosity determination using PCC method for Phase 2 HL-LHC can be based on counting the number of clusters or coincidences. The innermost ring of the last disk of TEPX (D4R1) is located at 2.65 m away from the interaction point that is beyond the tracking acceptance ($|\eta = 4|$) and as this region has few tracking points, it can be solely used for the purpose of luminosity measurement by using the full available trigger rate and bandwidth.  Two fold coincidences are those hits which are created by the module overlap regions between various layers of one TEPX double disk. Two fold coincidences are better way to distinguish between a real hit and random electrical noise. They are more likely to be real hit than random electrical noise. Two fold coincidence in $\phi$ will involve modules overlapping in the same ring in front and back layers of one double disk as shown in top part of Fig. 30 and Fig. 31 while two fold coincidence in r will require modules overlapping between successive rings in the front (L1 $\&$ L2) and back layers (L3 $\&$ L4) of one double disk as shown in bottom part of Fig. 30. Luminosity calculated based on counting coincidences has an advantage over clusters that afterglow effects are tiny in the case of coincidences. \\

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.8\columnwidth]{./tepx_geometry.png}
  \caption{ \onehalfspacing A layout of the CMS Phase-2 inner tracker showing four TEPX disks, eight TFPX disks and 4 barrel layers.}
  \label{fig:CMS}
\end{figure}

>>>>

The High Luminosity (HL)-LHC will increase instantaneous luminosity to unprecedented value of $7.5 \times 10^{34} cm^{-2} s^{-1}$ which corresponds to 200 proton-proton collisions per bunch crossing (pileup). Run 2 pixel detector will not be able to handle the extreme radiation environment, resolve nearby particle tracks and operate properly to give a reliable estimate of the instantaneous luminosity for high pileup values. That is why it will be replaced by a new pixel detector which will be composed of three subdetectors: tracker barrel pixel detector (TBPX), tracker forward pixel detector (TFPX) and tracker endcap pixel detector (TEPX). TEPX will have better radiation tolerance, increased granularity, improved two-track separation, improved estimation of hit rate and statistical precision, extended tracking acceptance $|\eta|=4$ with Disk 4 Ring 1 operating as an independent luminometer \cite{Klein:2017nke}. \\

Tracker endcap pixel detector (TEPX) consists of four double disks per side (-Z and +Z) as shown in Fig. 27 with each double disk containing five rings as shown in Fig. 28 having 20, 28, 36, 44 and 48 modules respectively. One double disk has four surfaces with +Z side containing modules with even module number in front layers (L1 $\&$ L2) and modules with odd module number in back layers (L3 $\&$ L4) from Ring 1 to Ring 4 and for Ring 5, modules with odd module number in front layers and modules with even module number in back layers as shown in Fig. 29. For -Z side, four surfaces contain modules with odd module number in front layer and modules with even module number in back layer from Ring 1 to Ring 4 and for Ring 5, modules with even module number in front layer and modules with odd module number in back layer . \\

Instantaneous luminosity determination using PCC method for Phase 2 HL-LHC can be based on counting the number of clusters or coincidences. The innermost ring of the last disk of TEPX (D4R1) is located at 2.65 m away from the interaction point that is beyond the tracking acceptance ($|\eta = 4|$) and as this region has few tracking points, it can be solely used for the purpose of luminosity measurement by using the full available trigger rate and bandwidth.  Two fold coincidences are those hits which are created by the module overlap regions between various layers of one TEPX double disk. Two fold coincidences are better way to distinguish between a real hit and random electrical noise. They are more likely to be real hit than random electrical noise. Two fold coincidence in $\phi$ will involve modules overlapping in the same ring in front and back layers of one double disk as shown in top part of Fig. 30 and Fig. 31 while two fold coincidence in r will require modules overlapping between successive rings in the front (L1 $\&$ L2) and back layers (L3 $\&$ L4) of one double disk as shown in bottom part of Fig. 30. Luminosity calculated based on counting coincidences has an advantage over clusters that afterglow effects are tiny in the case of coincidences.  \cite{Collaboration:2706512} \\

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.8\columnwidth]{./tepx_geometry.png}
  \caption{ \onehalfspacing A layout of the CMS Phase-2 inner tracker showing four TEPX disks, eight TFPX disks and 4 barrel layers. \cite{Klein:2017nke}}
  \label{fig:CMS}
\end{figure}

### 2021-05-17 project_ashish_monograph/vdM.tex
references added

The instantaneous luminosity when the crossing angle between the colliding bunches is negligible and bunches collide head on is given by the following integral, \\

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.7\columnwidth]{./vdmfit.png}
  \caption{ \onehalfspacing Example vdM scans for PCC for BCID 41, from the last scan pair in fill 4954, showing the rate normalized by the product of beam currents as a function of the beam separation $\Delta x$ and $\Delta y$ in the x (left) and y (right) direction, and the fitted curves. The purple curve shows the overall double-Gaussian fit, while the blue, yellow, and green curves show the first and second Gaussian components and the constant component, respectively. The lower panels display the difference between the measured and fitted values
divided by the statistical uncertainty\cite{}.}
  \label{fig:CMS}
\end{figure}

>>>>

The instantaneous luminosity when the crossing angle between the colliding bunches is negligible and bunches collide head on is given by the following integral, \cite{CMS-PAS-LUM-13-001} \\

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.7\columnwidth]{./vdmfit.png}
  \caption{ \onehalfspacing Example vdM scans for PCC for BCID 41, from the last scan pair in fill 4954, showing the rate normalized by the product of beam currents as a function of the beam separation $\Delta x$ and $\Delta y$ in the x (left) and y (right) direction, and the fitted curves. The purple curve shows the overall double-Gaussian fit, while the blue, yellow, and green curves show the first and second Gaussian components and the constant component, respectively. The lower panels display the difference between the measured and fitted values
divided by the statistical uncertainty\cite{CMS-PAS-LUM-15-001}.}
  \label{fig:CMS}
\end{figure}
